---
title: "{{< iconify line-md alert-circle >}} Info"
format: html
---

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
dashboardr::.page_config(accessibility = TRUE, crosstab_prefix = "info")
```

```{r setup}
#| echo: false
#| warning: false
#| message: false
#| error: false
#| results: 'hide'

# Load dashboardr (includes dplyr, highcharter as dependencies)
library(dashboardr)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)

```


## About This Dashboard

This dashboard tracks political advertising spending across three major Dutch data sources
leading up to the **2025 parliamentary elections**.

### Data Sources

**PolitiekeReclame.nl** - The central platform developed by five Dutch media associations
(Audify, MMA, NDP Nieuwsmedia, Outreach, and Screenforce) for EU political advertising
transparency. Political advertisers submit declarations covering TV, radio, newspapers,
magazines, outdoor, and online advertising.

**Ster** - The advertising sales organization for Dutch public broadcasting (NPO).
Their transparency page lists political parties and organizations that purchased
advertising through television, radio, and online channels.

**DPG Media** - A major Dutch media company operating newspapers, magazines, and
digital platforms. Their transparency page lists political advertising campaigns on
their print and digital channels. DPG Media reports spending boundaries rather than
exact amounts; we calculate spending by taking the midpoint.

### Methodology

- Data is fetched daily from all three APIs/sources
- Party names are aggregated using regex matching to unify different spellings
- For PolitiekeReclame channel analysis, multi-channel campaigns are split equally
- DPG Media spending is estimated as the midpoint of budget brackets
- Ster data is extracted from PDF transparency statements

### Code

Powered by the **reclamer** R package: [github.com/favstats/reclamer](https://github.com/favstats/reclamer)

*Dashboard generated: February 28, 2026 at 17:12 CET*

